Interpretable Machine Learning

### 2018-12-13 poster/poster_2018.Rnw
create talk for causality workshop;

\column{0.33}
\block{Machine Learning}{
Machine learning is a set of methods that allow computers to learn from data to make and improve predictions (e.g. cancer, weekly sales, credit default).
<<ml>>=
knitr::include_graphics("figure/programing-ml.png")
@
}
\block{Black Box Problem}{
<<iml>>=
knitr::include_graphics("figure/iml.png")
@

@@

\block{Why Explain?}{
We need to explain machine learning models to ... \cite{adadi2018peeking}:
% \begin{itemize}
% \item \textbf{Explain to justify}: Machine learning models make errors, can be unfair (e.g racial bias.) and make unexpected predictions. In those cases, we require a model to justify predictions.
% \item \textbf{Explain to control}: Understand more about the systems vulnerabilities and conditions for failure. e.g. know for what type of data it performs worse. 
% \item \textbf{Explain to improve}: Insights about model behaviour helps to improve the model.
% \item \textbf{Explain to discover}: A black box model is optimized to make predictions, not to give insights.
% Interpretability methods enable to extract insights from black box models.
% \end{itemize}
<<explain-to>>=
knitr::include_graphics("figure/explain.png")
@
}

@@

<<rent>>=
knitr::include_graphics("figure/rent-ml.png")
@

@@

ggplot(impres) + geom_col(aes(y = importance, x = feature)) + coord_flip() +
  theme(aspect.ratio=2/4, text = element_text(size=30),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_x_discrete("") +
  scale_y_continuous("Feature Importance")
@
\textit{In the rent example the most important feature was the size of the living area. Shuffling the size feature increases the model error by a factor of 2.5.}
}
<<apartment, echo=FALSE,include=FALSE>>=
# Choose apartment
apartment = data.frame(location = factor("Good", levels = c("Bad", "Good")),
    cats = factor("No", levels = c("No", "Yes")),
  size = 50)
pred.rent = round(pred$predict(apartment)$.prediction, 0)
@
\block{Individual Predictions}{
Sometimes we want to explain why a certain prediction was made by a machine learning model.
\textit{Suppose we want to explain the 989Euro rent prediction for a $50m^2$ apartment in a good location and where cats are forbidden.}
<<rent2, out.width="0.2\\textwidth">>=
knitr::include_graphics("figure/rent.png")
@

@@

%The goal is to find the best possible weights $\beta$.
<<lm>>=
knitr::include_graphics("figure/lm.png")
@
\textit{In the rent example, the following weights were learned:}
<<train-model-lm, echo=FALSE, include=FALSE>>=
mod.lm = lm(rent ~ ., data = df)
cf = round(coef(mod.lm))
knitr::kable(cf, format = 'latex')
@

@@

<<train-model-tree, echo=FALSE, out.width="0.28\\textwidth">>=
library(partykit)
mod.tree = ctree(rent ~ ., data = df, control = ctree_control(maxdepth = 2))
#plot(mod.tree, inner_panel = node_inner(mod.tree, pval = FALSE), type='simple')
knitr::include_graphics("figure/tree.png")
@
\textit{The tree predicts a rent of 1908 Euros for an apartment larger than 77 $m^2$ in a good location.}
\\ \\
\textbf{Decision Rules} \\
Decision rules are IF-THEN statements that consist of a condition and a prediction.
One or more rules can be used to make predictions. For example:
<<train-model-rules, echo=FALSE, includee=FALSE, eval=FALSE>>=

@@

}
\block{Surrogate Models}{
\textbf{Global Surrogate Models} \\
%Flexible but less interpretable models often predict the data better than interpretable models.
Intrinsically interpretable models (e.g. a tree) can be used to increase interpretability of a black box model (e.g. a neural network) by approximating its predictions by acting as a surrogate model that can be interpreted. 
<<surrogate>>=
knitr::include_graphics("figure/global-surrogate.png")
@

\textbf{Local Surrogate Models}\cite{ribeiro2016should} \\
To explain an individual prediction, replace a  complex model with a locally weighted, interpretable model (e.g. decision tree).
Data points that are close to the data point of interest get a high weight.
\vspace{0.5cm}
<<lime, echo=FALSE>>=
# locmod = LocalModel$new(pred, k = 2)
# locmod$explain(df[1,1:3])
# plot(locmod)
knitr::include_graphics("figure/local-surrogate.png")
@
}

>>>>

\column{0.33}
\block{Machine Learning}{
Machine learning is a set of methods that allow computers to learn from data to make and improve predictions (e.g. cancer, weekly sales, credit default).
<<ml>>=
knitr::include_graphics("../images/programing-ml.png")
@
}
\block{Black Box Problem}{
<<iml>>=
knitr::include_graphics("../images/iml.png")
@

@@

\block{Why Explain?}{
We need to explain machine learning models to ... \cite{adadi2018peeking}:
% \begin{itemize}
% \item \textbf{Explain to justify}: Machine learning models make errors, can be unfair (e.g racial bias.) and make unexpected predictions. In those cases, we require a model to justify predictions.
% \item \textbf{Explain to control}: Understand more about the systems vulnerabilities and conditions for failure. e.g. know for what type of data it performs worse. 
% \item \textbf{Explain to improve}: Insights about model behaviour helps to improve the model.
% \item \textbf{Explain to discover}: A black box model is optimized to make predictions, not to give insights.
% Interpretability methods enable to extract insights from black box models.
% \end{itemize}
<<explain-to>>=
knitr::include_graphics("../images/explain.png")
@
}

@@

<<rent>>=
knitr::include_graphics("../images/rent-ml.png")
@

@@

ggplot(impres) + geom_col(aes(y = importance, x = feature)) + coord_flip() +
  theme(aspect.ratio=2/4, text = element_text(size=30),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_x_discrete("") +
  scale_y_continuous("Feature Importance")
@
\textit{In the rent example the most important feature was the size of the living area. Shuffling the size feature increases the model error by a factor of 2.5.}
}
<<apartment, echo=FALSE,include=FALSE>>=
# Choose apartment
apartment = data.frame(location = factor("Good", levels = c("Bad", "Good")),
    cats = factor("No", levels = c("No", "Yes")),
  size = 50)
pred.rent = round(pred$predict(apartment)$.prediction, 0)
@
\block{Individual Predictions}{
Sometimes we want to explain why a certain prediction was made by a machine learning model.
\textit{Suppose we want to explain the 989Euro rent prediction for a $50m^2$ apartment in a good location and where cats are forbidden.}
<<rent2, out.width="0.2\\textwidth">>=
knitr::include_graphics("../images/rent.png")
@

@@

%The goal is to find the best possible weights $\beta$.
<<lm>>=
knitr::include_graphics("../images/lm.png")
@
\textit{In the rent example, the following weights were learned:}
<<train-model-lm, echo=FALSE, include=FALSE>>=
mod.lm = lm(rent ~ ., data = df)
cf = round(coef(mod.lm))
knitr::kable(cf, format = 'latex')
@

@@

<<train-model-tree, echo=FALSE, out.width="0.28\\textwidth">>=
library(partykit)
mod.tree = ctree(rent ~ ., data = df, control = ctree_control(maxdepth = 2))
#plot(mod.tree, inner_panel = node_inner(mod.tree, pval = FALSE), type='simple')
knitr::include_graphics("../images/tree.png")
@
\textit{The tree predicts a rent of 1908 Euros for an apartment larger than 77 $m^2$ in a good location.}
\\ \\
\textbf{Decision Rules} \\
Decision rules are IF-THEN statements that consist of a condition and a prediction.
One or more rules can be used to make predictions. For example:
<<train-model-rules, echo=FALSE, includee=FALSE, eval=FALSE>>=

@@

}
\block{Surrogate Models}{
\textbf{Global Surrogate Models} \\
%Flexible but less interpretable models often predict the data better than interpretable models.
Intrinsically interpretable models (e.g. a tree) can be used to increase interpretability of a black box model (e.g. a neural network) by approximating its predictions by acting as a surrogate model that can be interpreted. 
<<surrogate>>=
knitr::include_graphics("../images/global-surrogate.png")
@

\textbf{Local Surrogate Models}\cite{ribeiro2016should} \\
To explain an individual prediction, replace a  complex model with a locally weighted, interpretable model (e.g. decision tree).
Data points that are close to the data point of interest get a high weight.
\vspace{0.5cm}
<<lime, echo=FALSE>>=
# locmod = LocalModel$new(pred, k = 2)
# locmod$explain(df[1,1:3])
# plot(locmod)
knitr::include_graphics("../images/local-surrogate.png")
@
}
